### 2015-12-15 untitled.tex
restructure background, clarify definitions

We briefly introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then explain the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.

\subsection{Preliminaries}
\label{sec:prelim}

@@

In contrast, the segmented sporadic task model extends the above four-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals. These intervals are represented as a tuple
$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals. For the simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment. The arguments can be easily extended to handle tasks that start with a self-suspension.

@@

@@

	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem. Recent findings by Nelissen et al. \cite{ecrts15nelissen} can be applied, but their method takes exponential time.
\end{enumerate}
%
Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}): while it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention, it is actually unsafe to do so.
\begin{enumerate}
\setcounter{enumi}{2}
	\item The period enforcer algorithm invalidates all existing blocking analyses as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
\end{enumerate}

@@

Next, let us consider the same task set under control of the period enforcer algorithm.
The period enforcer algorithm determines for each segment (i.e., for each converted single-segment deferrable task) an \emph{eligibility time}. If a segment resumes (i.e., a job of a single-segment deferrable task arrives) before its eligibility time, the execution of the segment (i.e., the single-segment job) is delayed until the eligibility time is reached.

A segment's eligibility time is determined according to the following rule. Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth segment of the $j$\xth job of task $\tau_i$. Further, let $a^k_{i,j}$ denote the segment's arrival time.\footnote{A segment \emph{arrives} when it becomes available for execution. The first segment arrives immediately when the job is released; the second segment arrives when the job resumes from its self-suspension.} Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$. According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
\begin{align}\label{eq:ET-def}
	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
\end{align}
where $ET_{i,0}^k = -T_i$.

Assuming this definition, Figure \ref{fig:example} shows the resulting schedule for a periodic release pattern. The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible. Note that the first segment of $\tau_2$'s first job is delayed due to interference from $\tau_1$. As a result, the second segment of $\tau_2$'s first job does not resume until time $a^2_{2,1} = 9$. Thus, we have
\begin{align*}
	ET_{2,1}^1 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 0)\right) = 0  \text{ and }
\\
	ET_{2,1}^2 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 9\right) ) = 9.
\end{align*}

>>>>

We briefly introduce the needed background in Section~\ref{sec:prelim}, restate our contributions more precisely in Section~\ref{sec:questions}, and then establish the three above  observations in detail in Sections \ref{sec:unschedulable}--\ref{sec:locking} before concluding in Section \ref{sec:conclusion}.

\section{Preliminaries}
\label{sec:prelim}

The period enforcer algorithm applies to self-suspending tasks on uniprocessors under fixed-priority scheduling, and hence by extension also to multiprocessors under partitioned fixed-priority scheduling (where tasks are statically assigned to processors and each processor is scheduled as a uniprocessor). In this section, we review the underlying task model (Section \ref{sec:taskmodel}), introduce the period enforcer algorithm (Section \ref{sec:pe}), summarize its analysis (Section \ref{sec:classic-analysis}), and finally restate our observations (Section \ref{sec:questions}).

\subsection{Self-Suspending Tasks}
\label{sec:taskmodel}

@@

In contrast, the segmented sporadic task model extends the above four-tuple by characterizing each self-suspending task as a (fixed) finite linear sequence of computation and suspension intervals. These intervals are represented as a tuple
$(C_{i}^1,S_{i}^1,C_{i}^2,S_{i}^2,...,S_{i}^{m_i-1},C_{i}^{m_i})$, which is composed of $m_i$ computation segments separated by $m_i-1$ suspension intervals. For simplicity of presentation, we assume that a task $\tau_i$ always starts with a computation segment. The arguments can be easily extended to handle tasks that start with a self-suspension.

@@

We say that a segment \emph{arrives} when it becomes available for execution. The first computation segment arrives immediately when the job is released; the second computation segment (if any) arrives when the job resumes from its first self-suspension, \textit{etc.} For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task). A \emph{level-$i$ busy interval} is a maximal interval during which  the processor executes only segments of tasks with priority $i$ or higher.

\subsection{The Period Enforcer Algorithm}
\label{sec:pe}

The key idea underlying the period enforcer algorithm is to artificially delay the execution of computation segments if a job resumes ``too soon.'' To this end,  the period enforcer algorithm determines for each computation segment an \emph{eligibility time}. If a segment resumes  before its eligibility time, the execution of the segment is delayed until the eligibility time is reached.

A segment's eligibility time is determined according to the following rule. Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth computation segment of the $j$\xth job of task $\tau_i$. Further, let $a^k_{i,j}$ denote the segment's arrival time. Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$). According to Section 3.1 in \cite{Raj:suspension1991}, the period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
\begin{align}\label{eq:ET-def}
	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
\end{align}
where $ET_{i,0}^k = -T_i$.

% Assuming this definition,

\subsection{Classic Analysis of the Period Enforcer Algorithm}
\label{sec:classic-analysis}

@@

	\item Deriving a single-segment deferrable task set corresponding to a given set of sporadic segmented self-suspending tasks in polynomial time is an open problem. Recent findings by Nelissen et al. \cite{ecrts15nelissen} can be applied, but their method takes exponential time.
\end{enumerate}
%
Finally, we make a third observation concerning the use of the period enforcer in conjunction with suspension-based multiprocessor locking protocols for partitioned fixed-priority scheduling (such as the MPCP~\cite{LNR:09,Ra:90} or the FMLP~\cite{BLBA:07,BA:08}): while it is certainly tempting to apply period enforcement with the intention of avoiding the negative effects of deferred execution due to lock contention, it is actually unsafe to do so.
\begin{enumerate}
\setcounter{enumi}{2}
	\item The period enforcer algorithm invalidates all existing blocking analyses as there exist non-trivial feedback cycles between the period enforcer rules and blocking durations.
\end{enumerate}

@@

Next, let us consider the same task set under control of the period enforcer algorithm, as defined in Section \ref{sec:pe}.
Figure \ref{fig:example} shows the resulting schedule for a periodic release pattern. The first job of task $\tau_2$ (which arrives at time $a^1_{2,1} =  0$) is executed as if there is no period enforcement since the definition $ET_{2,0}^1 = ET_{2,0}^2 = -T_2$ ensures that both segments are immediately eligible. Note that the first segment of $\tau_2$'s first job is delayed due to interference from $\tau_1$. As a result, the second segment of $\tau_2$'s first job does not resume until time $a^2_{2,1} = 9$. Thus, we have
\begin{align*}
	ET_{2,1}^1 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 0)\right) = 0  \text{ and }
\\
	ET_{2,1}^2 & = \max\left(-T_2 + T_2,\ \mathit{busy}(\tau_2, 9\right) ) = 9.
\end{align*}
